---
output:
  html_document: default
  pdf_document: default
---
Red Wine quality by Alicia Escontrela
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(gridExtra)
library(dplyr)
library(RColorBrewer)
library(reshape2)
```

```{r echo=FALSE, message=FALSE}
# Load the Data
wine <- read.csv('wineQualityWhites.csv')
str(wine)
```
```{r echo=FALSE, message=FALSE}
#View data
summary(wine)
```


# Univariate Plots Section


```{r echo=FALSE, warning=FALSE, message=FALSE}
# Function to plot a histogram changing variable and bins
one_var_plot<-function (var1, bin_val, nbins)
{
  ggplot(aes(x = wine[,var1]), data = wine) +
  geom_histogram(binwidth = bin_val, bins = nbins)+
  xlab(var1)+
  ylab('Frequency')
}

one_var_plot('quality', 0.5, 10)
```

> Most of the wine anlysed on this dataset have medium quality. In order to 
visualize this variable in context I'll create a categorical variable named 
quality_factor, using this ranges: 0-4 <- Low, 5-6 <- Medium, 
7-9 <- High 


```{r echo=FALSE, warning=FALSE, message=FALSE}
#Create a categorical variable for quality
wine$quality_factor <- cut(wine$quality, breaks = c(0,4,6,9), 
                           labels = c("Low", "Medium", "High"))
names(wine)
ggplot(aes(x = quality_factor), data = wine) +
  geom_bar(stat="count")
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
#Plots for fixed acidity

h1 <- one_var_plot('fixed.acidity', 0.1, 10) +
  scale_x_continuous(limits = c(4,14))+
  labs(y = 'Fixed Acidity (g / dm^3)', x = '',
          title = "Fixed acidity distribution") 

bp1 <- ggplot(wine, aes(y = fixed.acidity, x = 1)) +
  geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = fixed.acidity), alpha = 0.05) +
  coord_flip()+
  labs(y = 'Fixed Acidity (g / dm^3)', x = '',
          title = "Fixed acidity distribution") 

d1 <- ggplot(aes(x = fixed.acidity), data = wine) +
 geom_density()+
 labs(x = 'Fixed Acidity (g / dm^3)',
          title = "Fixed acidity distribution") 

  grid.arrange(h1, bp1, d1)

```
  
  
> Plots above shows that fixed acidity has a positive skewed distribution. 
It would be interesting to compare this variables with other types of
acidity, as we will se below

```{r echo=FALSE, warning=FALSE, message=FALSE}  
h2 <- one_var_plot('volatile.acidity', 0.05, 10) + 
  ggtitle('Volatile Histogram')+
  scale_x_continuous()+
  labs(y = 'Volatile Acidity (g / dm^3)', x = '',
          title = "Volatile acidity distribution") 

bp2 <- ggplot(wine, aes(y = volatile.acidity, x = 1)) +
  geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = volatile.acidity), alpha = 0.05) +
  coord_flip()+
  labs(y = 'Volatile Acidity (g / dm^3)', x = '',
          title = "Volatile acidity distribution") 

d2 <- ggplot(aes(x = volatile.acidity), data = wine) +
 geom_density()+
 labs(x = 'Volatile Acidity (g / dm^3)',
          title = "Volatile acidity distribution") 

 grid.arrange(h2, bp2, d2)
 
```
 

> Plot above shows that volatile acidity variable has a positive skewed 
distribution; where there are a lot of outliers, as we can see on 
boxplot

```{r echo=FALSE, warning=FALSE, message=FALSE}  
h3 <- one_var_plot('citric.acid', 0.025, 10) + 
  ggtitle('Citric acid distribution')+
  scale_x_continuous(limits = c(0.1,1.66))


bp3 <- ggplot(wine, aes(y = volatile.acidity, x = 1)) +
  geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = volatile.acidity), alpha = 0.05) +
  coord_flip()+
  labs(y = 'Citric acid (g / dm^3)', x = '',
          title = "Citric acid distribution") 

d3 <- ggplot(aes(x = citric.acid), data = wine) +
 geom_density()+
 labs(x = 'Citric acid (g / dm^3)',
          title = "Citric acid distribution") 

grid.arrange(h3, bp3, d3)

```


> Citric acid ditribution plots shows a distribution more similar to normal 
than fixed and volatile acidity and that it has a lot of outliers.
Also, is important to notice that x limits are similar for volatile 
acidity and citric acid (0-1), while data is more widespread than fixed 
acidity. 



```{r echo=FALSE, warning=FALSE, message=FALSE}
#Plots for variables related with sulfure.dioxide
h4 <- one_var_plot('free.sulfur.dioxide', 1, 50)+
  ggtitle('Free Sulfur Dioxide Histogram')+
  scale_x_continuous(limits = c(0,100))+
  labs(y = 'free.sulfur.dioxide (mg / dm^3)', x = '',
          title = "Free Sulfur Dioxide distribution") 

bp4 <- ggplot(wine, aes(y = free.sulfur.dioxide, x = 1)) +
  geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = free.sulfur.dioxide), alpha = 0.03) +
  coord_flip()+
  labs(y = 'free.sulfur.dioxide (mg / dm^3)', x = '',
          title = "Free Sulfur Dioxide distribution") 

d4 <- ggplot(aes(x = volatile.acidity), data = wine) +
 geom_density()+
 labs(x = 'free.sulfur.dioxide (mg / dm^3)', x = '',
          title = "Free Sulfur Dioxide distribution")

 grid.arrange(h4, bp4, d4)
 
```


> Free sulfure dioxide has a positive skewed distribution, that has 
a lot of outliers, as we can see on boxplot

```{r echo=FALSE, warning=FALSE, message=FALSE}

h5 <- one_var_plot('total.sulfur.dioxide', 5, 50) +
  ggtitle('Total Sulfur Dioxide distribution')+
  scale_x_continuous(limits = c(0,300))

bp5 <- ggplot(wine, aes(y = total.sulfur.dioxide, x = 1)) +
  geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = total.sulfur.dioxide), alpha = 0.05) +
  coord_flip()+
  labs(y = 'total.sulfur.dioxidey (mg / dm^3)', x = '',
          title = "Total Sulfur Dioxide distribution") 

d5 <- ggplot(aes(x = total.sulfur.dioxide), data = wine) +
 geom_density()+
 labs(x = 'total.sulfur.dioxide (mg / dm^3)', x = '',
          title = "Total Sulfur Dioxide distribution")

grid.arrange(h5, bp5, d5)
  
```


> Total sulfure dioxide has a positive skewed distribution, 
with less outliers than free sulfure dioxide.


> Since free sulfur dioxide prevents microbial growth and 
oxidation of wine and total sulfur dioxide becomes evident in the nose 
and taste of the wine, as mentioned on description of variables, it would
be interesting to evaluate their impact in quality rate of wine

```{r echo=FALSE, warning=FALSE, message=FALSE}

h6 <- one_var_plot('sulphates', 0.01, 10)+
  ggtitle('Sulphates distribution')+
  scale_x_continuous()

bp6 <- ggplot(wine, aes(y = sulphates, x = 1)) +
  geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = sulphates), alpha = 0.05) +
  coord_flip()+
  labs(y = 'sulphates (g / dm3)', x = '',
          title = "Sulphates distribution") 

d6 <- ggplot(aes(x = sulphates), data = wine) +
 geom_density()+
 labs(y = 'sulphates (g / dm3)', x = '',
          title = "Sulphates distribution") 

grid.arrange(h6, bp6, d6)

```


> Plots above show that sulphates present on this dataset has a 
positive skewed distribution, with a strong presence of outliers

```{r echo=FALSE, warning=FALSE, message=FALSE}

h7 <- one_var_plot('alcohol', 0.1, 50)+
   ggtitle('Alcohol distribution')+
   scale_x_continuous()


bp7 <- ggplot(wine, aes(y = alcohol, x = 1)) +
  geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = alcohol), alpha = 0.05) +
  coord_flip()+
  labs(x = 'alcohol (% by volume)', x = '',
          title = "Alcohol distribution")

d7 <- ggplot(aes(x = alcohol), data = wine) +
 geom_density()+
 labs(x = 'alcohol (% by volume)', x = '',
          title = "Alcohol distribution")


grid.arrange(h7, bp7, d7)

```


> % of alcohol by volume shows a positive skewed distribution that is
more widespread than other variables

```{r echo=FALSE, warning=FALSE, message=FALSE}


h8 <- one_var_plot('pH', 0.05, 5)+
   ggtitle('pH distribution')+
   scale_x_continuous()


bp8 <- ggplot(wine, aes(y = pH, x = 1)) +
  geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = pH), alpha = 0.05) +
  coord_flip()+
  labs(y = 'pH', x = '',
          title = "pH distribution") 

d8 <- ggplot(aes(x = pH), data = wine) +
 geom_density()+
 labs(x = 'pH', x = '',
          title = "pH distribution")

grid.arrange(h8, bp8, d8)
  
```


> Plots above show that sulphates and alcohol have a positive skewed 
distribution, while pH has a distribution more similar to normal. Also, as 
states on pH description, you can notice that pH for most of wines on the 
dataset are between 3 and 4 


```{r echo=FALSE, warning=FALSE, message=FALSE}
h9 <- one_var_plot('residual.sugar', 0.3, 5)+
   scale_x_continuous()+
  ggtitle("Residual sugar distribution")

bp9 <- ggplot(wine, aes(y = residual.sugar, x = 1)) +
  geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = residual.sugar), alpha = 0.05) +
   coord_flip()+
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)+
  labs(y = 'residual.sugar (g / dm^3)', x = '',
          title = "Residual sugar distribution")


d9 <- ggplot(aes(x = residual.sugar), data = wine) +
 geom_density()+
 labs(x = 'residual.sugar (g / dm^3)', x = '',
          title = "Residual sugar distribution")

grid.arrange(h9, bp9, d9)

```


> Plots above show that residual sugar have a positive skewed distribution, 
with almost none outliers. Also, we can notice that there are more 
concentration of wines within first quantile and resiudal sugar and
almost none above 20 g / dm^3. 


```{r echo=FALSE, warning=FALSE, message=FALSE}


bp10 <- ggplot(wine, aes(y = density, x = 1)) +
  geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = density), alpha = 0.05) +
  coord_flip()+
  labs(y = 'Density', x = '',
          title = "Density distribution") 

d10 <- ggplot(aes(x = density), data = wine) +
 geom_density()+
 labs(x = 'Density', x = '',
          title = "Density distribution")

grid.arrange(bp10, d10)
```


> Plots above show that most of wines in the dataset have a density
between 0.99 and 1. We didn't find usefull to plot a histogram
to this variable because we haven't found a resolution good enough


# Univariate Analysis

### What is the structure of your dataset?
There are 4898 observations of  13 variables (fixed.acidity, volatile.acidity, 
citric.acid, residual.sugar, chlorides, free.sulfur.dioxide, 
total.sulfur.dioxide, density, pH, sulphates, alcohol, quality); 11 of these 
variables describe chemicals properties and there is one variable that describe 
overall quality (0 as bad quality and 10 as high quality).

Most of the variables show a positive skewed distribution. 
However these variables shows a distribution more similar to normal:
pH, density. 

In the first exploratory analysis, we can notice that most of wines evaluated 
have rated as medium quality with a density very similar to each other 
(the difference between min and max values is about 0.052), pH from 3 to 4, 
residual sugar less than 20 g/dm^3 with a mean value of 6.391  g/dm^3

### What is/are the main feature(s) of interest in your dataset?
The main feauture of interest in the dataset is quality. We are interested to 
evaluated how other chemical properties have an impact on quality rating

### What other features in the dataset do you think will help support your 
investigation into your feature(s) of interest? 
We think chemicals properties such as acidity, pH and residual sugar would have
an impact on the quality of the wine, because these variables have impact on 
wine taste

### Did you create any new variables from existing variables in the dataset?
Yes, we created one categorical variable for quality because we found that 
would be more usefull to visualize giving to that variable a categorical context

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
No, we just create one new categorical variable for quality

# Bivariate Plots Section

> I'm interested to evaluate the relation between quality and other chemicals
properties, to make anlaysis more simple I've created a function to graph 
boxplot to compare quality with another variable

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Function to create a boxplot to evaluate quality_factor with other variables
quality_2var_boxplot<-function (var)
{
  ggplot(aes( x= quality_factor, y = wine[,var]), data = wine) +
  geom_boxplot(aes(group = quality_factor))+
    ylab(var)+
  xlab('Quality')
}

```


> We are going to start bivariate analysis by measuring correlation between
variables using ggpair and ggcor as shown below


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=12}

ggcorr(wine[, 2:13], geom = "blank", label = TRUE, hjust = 0.6) +
  geom_point(size = 10, aes(color = coefficient > 0,
                            alpha = abs(coefficient) > 0.45)) +
  scale_alpha_manual(values = c("TRUE" = 0.25, "FALSE" = 0)) +
  guides(color = FALSE, alpha = FALSE)

  
```



> After first correlation analysis you can see that variables with 
stronger correlation are 
density/alcohol and density/residual sugar
density vs alcohol (-0.8)
density vs residual.sugar (0.8)
Following variables show correlation around 0.4 - 0.6
total.sulfur.dioxide vs free.sulfur.dioxide (0.6)
total.sulfur.dioxide vs density (0.5)
alcohol vs quality (0.4)
alcohol vs residual sugar (-0.5)After first bivariate analysis and found variables wit most correlation,


> So, we subset the dataset for these variables and appy ggpair

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height12}

set.seed(2500)

wine_subset <- subset(wine, select= c('density', 'quality', 
                                      'alcohol', 'residual.sugar', 
                                      'total.sulfur.dioxide', 
                                      'free.sulfur.dioxide'))

wine_sample <- wine_subset[sample(1:length(wine$quality), 1000),]
ggpairs(wine_sample, 
        lower = list(continuous = wrap("points", shape = I('.'))),
        upper = list(combo = wrap("box", outlier.shape = I('.'))))+
  theme(axis.text = element_blank())

  
ggsave("plotpairs_subset.png")
```




> Using function created above, we are going to analyse each variable
in relation with quality

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Quality - acidity evaluation

box_q_fixed_ac <- quality_2var_boxplot('fixed.acidity')
box_q_volatile_ac <- quality_2var_boxplot('volatile.acidity')
box_q_citric_ac <- quality_2var_boxplot('citric.acid')

grid.arrange(box_q_fixed_ac, box_q_volatile_ac, box_q_citric_ac, ncol=3)
```

> Is important to notice that there are a lot of outliers for volatile acidity 
and citric acid. 
Regarding  citric.acid; graph above show that values are similar for all range 
of qualities. However, for low quality wines the difference between first and 
third quantile are bigger that for high quality wines
Volatile acidity boxplot shows that low quality wines have higher level of this 
variable, which can lead to vinegar taste 
Fixed acidity boxplot shows a similar behaviour. We should remove outliers for
a deep analysis


```{r echo=FALSE, warning=FALSE, message=FALSE}
#Boxplot fixed acidity removing outliers

box_q_fixed_ac2 <- quality_2var_boxplot('fixed.acidity')+
  ylim(c(5,9))+
  ggtitle('Fixed acidity - quality (removing outliers)')

box_q_fixed_ac2 
```


> Graph above shows fixed acidity boxplot group by quality_factor variable 
removing outliers shown before

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Quality - sugar and pH boxplot removing outliers

box_q_sugar <- quality_2var_boxplot('residual.sugar') + ylim(c(0,15)) 
box_q_pH <- quality_2var_boxplot('pH') + ylim(c(2.5, 3.5))

grid.arrange(box_q_sugar, box_q_pH, ncol=2)
```


> pH graph show that all wines in dataset have similars pH between 3-4 with 
similar behaviuor. 
Regarding residual sugar, graph shows that medium quality wines have higher 
concentration of residual sugar than high quality.


```{r echo=FALSE, warning=FALSE, message=FALSE}
#Quality - density and alcohol
box_q_density <- quality_2var_boxplot('density') + ylim(c(0.98, 1))
box_q_alcohol <- quality_2var_boxplot('alcohol') 

grid.arrange(box_q_density, box_q_alcohol, ncol=2)
```


> Graphs above show that relationship between density and alcohol seems to be 
inversely proportional; which means that when density increases concentration of 
alcohol decreases. For instante, high quality wine have lower density but higher 
percent of alcohol, while low quality wine have higher density but lower 
percentange of alcohol.


```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(aes(x= density, colour=quality_factor), data = wine) +
  geom_density()+
  scale_color_brewer(type='seq') +
  xlim(c(0.9871,1.0390))+
  xlab('Density(g / cm^3)')
```


> Plot above for density show that behaviour is pretty similar to each other 
for low and medium quality, while density for high quality wine tend to be lower

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(aes( x= quality_factor, y = free.sulfur.dioxide/total.sulfur.dioxide), 
       data = wine) +
  geom_boxplot(alpha = .05,color = 'blue', aes(group = quality_factor))+
  geom_jitter( alpha = .3)  +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```

> Boxplot above shows that there is more concetration of SO2 for medium 
quality wines



```{r echo=FALSE, warning=FALSE, message=FALSE}
#Function to graph scatter plot for 2 variables
bivar_scatterplot<-function (xvar, yvar)
{
  ggplot(aes( x= wine[,xvar], y = wine[,yvar]), data = wine) +
  geom_point(alpha = 0.05, position = position_jitter(h=0)) +
    ylab(yvar) +
    xlab(xvar)
}
```



>  In order to evaluate how acid and residual sugar have an impact on quality 
rating  we found that there is a scale created by International Riesling 
Foundation (http://drinkriesling.com/tasteprofile/thescale) to categorise taste 
of wine. As mentioned on their webpage this categories should be used by 
wineries. 

> The IRF scale is based on sugar to acid ratio and then is corrected by pH. 
We create a new variable calle total acid that sum up all types of acid 
(wine$fixed.acidity + wine$volatile.acidity + wine$citric.acid) and the used 
this variable to calculate sugar to acid ratio

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Create a new variable for sugar to acid ratio
wine$total_acid<- wine$fixed.acidity + wine$volatile.acidity + wine$citric.acid
head(wine$total_acid)

wine$sugar_acid_ratio <- wine$residual.sugar/wine$total_acid
summary(wine$sugar_acid_ratio)

```

> Then we create a new categorical variable based on sugar_acid_ratio variable 
and the IRF scale. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Create a categorical variable for sugar_acid_ratio
wine$ratio_scale <- cut(wine$sugar_acid_ratio, breaks = c(0,1,2,4,8), 
                       labels = c("Dry", "Medium Dry", "Medium Sweet", "Sweet"))


summary(wine$ratio_scale)
```

> Then we created a function using variables ratio_scale and pH to measure the 
IRF scale taking into account how pH modified ratio_scale variable creted above

```{r echo=FALSE, warning=FALSE, message=FALSE}

scale_IRF <- function(ratio_scale, pH) {
    if(ratio_scale=="Dry") {
     if ((pH >=3.3) & (pH <=3.5)){
       "Medium Dry"
     }
     else if (pH > 3.5){
        "Medium Sweet"
     }
      else{
        "Dry"
      }
    }
   
  else if(ratio_scale=="Medium Dry") {
     if (pH >=3.3){
       "Medium Sweet"
     }
     else if (pH <= 2.9){
        "Dry"
     }
      else{
        "Medium Dry"
      }
    }
  
  else if(ratio_scale=="Medium Sweet") {
     if (pH >=3.3){
       "Sweet"
     }
     else if ((pH <= 2.9) & (pH > 2.8)){
        "Medium Dry"
     }
    else if (pH <= 2.8){
        "Dry"
     }
      else{
        "Medium Sweet"
      }
  }

  else if(ratio_scale=="Sweet") {
     if ((pH <= 2.9) & (pH > 2.8)){
        "Medium Sweet"
     }
    else if (pH <= 2.8){
        "Medium Dry"
     }
      else{
        "Sweet"
      }
  }
  
  else{
    "N/A"
  }
}


wine$scale <- mapply(scale_IRF, wine$ratio_scale, wine$pH)

p1 <- ggplot(aes(x=ratio_scale), data = wine)+
  geom_bar()

p2 <- ggplot(aes(x=scale), data = wine)+
  geom_bar()

grid.arrange(p1, p2, ncol =2)
```


> Plots above show how the category of wine have changed taking into account pH 
impact in taste; which show that IRF scale changed after adding pH impact on 
taste


```{r echo=FALSE, warning=FALSE, message=FALSE}
#Plot alcohol - density 
bivar_scatterplot('alcohol', 'density') +
  coord_trans(y="sqrt") +
  ylim(c(0.98 , 1.01))+
  geom_smooth(method = 'lm', color='red')
```

> Graph above shows that there are less density for a higher percentage of 
alcohol. The graph shows a lineal pattern for the relationship with the most 
strongest correlation evaluated on ggpairs

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(aes(x= alcohol, colour=quality_factor), data = wine) +
  geom_density()+
  scale_color_brewer(type='seq', palette = 2) +
  xlab('Alcohol (% by volume)')

```

> Alcohol is the variable with strongest correlation directly related with 
quality. Graph above shows that percentange of alcohol increases for high 
quality wines, while percentage of alcohol for medium and low quality wines are 
pretty similar 

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Plot density - residual sugar 
bivar_scatterplot('residual.sugar', 'density')+
  coord_cartesian(ylim = c(.986, 1.005), xlim = c(0,30))+
  geom_smooth(method = lm, se = FALSE)
  
```

> The relationship between residual sugar and density shows that there are less 
concentration of residual sugar for wine with high density. However the y limits 
are not wide spread, as seen before the limits of density are between 0.9871 and 
1.0390

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Plot Total sulfure dioxide - density
bivar_scatterplot('total.sulfur.dioxide','density') +
  scale_x_continuous(lim = c(0, quantile(wine$total.sulfur.dioxide, 0.99))) +
  scale_y_continuous(lim = c(0.9871, 1.0390))+
  stat_smooth(method = 'lm', color='red')
 
  


```

> Plot above shows that wines with more concentration of total sulfure dioxide 
tends to have more density

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Plot Total sulfure dioxide 

ggplot(aes( y= free.sulfur.dioxide, x = total.sulfur.dioxide), data = wine)+ 
  geom_point(alpha = 0.05, position = position_jitter(h=0))+
  geom_smooth(method = 'loess', color='purple')+
  coord_trans('log10')
  
```


> The relationship between free sulfure dioxide and total sulfure dioxide seems 
to be log10 with an increase on the total sulfure dioxide for more 
concrentration of free sulfure dioxide




```{r echo=FALSE, warning=FALSE, message=FALSE}
#Plot alcohol - residual sugar
bivar_scatterplot('alcohol','residual.sugar')+
  geom_smooth(method = 'loess', color='green')+
  ylim(c(0,25))
  
```


> The relation between alcohol and residual sugar are not so strong. The graph 
shows a tendency to decrease the concentration of alcohol for wines with less 
residual sugar. It would be interesting to evaluate this variables including 
quality and IRF scale for multivariate analysis 

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(aes( x= scale, y = alcohol), data = wine) +
  geom_boxplot(aes(group = scale))


```

> Plot above shows that there is more concentration of alcohol for Dry wines 
than medium sweet and sweet wines. It's important to notice that the alcohol 
concentration for dry and medium dry wines have a similar behaviour

```{r echo=FALSE, warning=FALSE, message=FALSE}
bivar_scatterplot('fixed.acidity', 'pH')+
  coord_trans('sqrt')+
  ylim(c(2.5,4))+
  xlim(c(5,10))+
  geom_smooth(method = 'loess', color='green')
  
```


> This graph shows a tendency of decrease pH for more concentration of fixed 
acidity; which makes sense because pH more acid is 0. However their correlation 
is not so strong, which is also seen in this graph

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(aes( x= sulphates, y = pH), data = wine) +
  geom_point(alpha = 0.05, position = position_jitter(h=0))+
  coord_trans('sqrt')+
  facet_wrap(~quality_factor)+
  coord_trans('sqrt')


```

> Graph above shows that there is more concentration on sulphates for medium 
quality wines 

> We're going to evaluated the relationship between variable quality_factor and 
alcohol based on conditional means, as seen below


```{r echo=FALSE, warning=FALSE, message=FALSE}
#Plot alcohol - quality

  
quality_groups <- group_by(wine, quality)
wine_by_quality <- summarise(quality_groups,
                          al_mean = mean(alcohol),
                          al_median = median(alcohol),
                          n=n())
wine_by_quality <- arrange(wine_by_quality, quality)

wine_by_quality

ggplot(aes(x=quality, y=al_mean), data=wine_by_quality) + 
  geom_bar(stat = 'summary', fun.y = mean, color = "blue") 
  
 
  
```


> Plot above shows that mean concentration of alcohol are bigger for high 
quality wines. Wines evaluated with quality of 5 are the ones with less 
concentration of alcochol, which means less alcohol than low quality wines


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Ggpairs analysis give us that the variable that have more strong correlation 
with quality is alcohol (0.482). Exploratory analysis regarding these variables
show that high quality wines have more percentage of alcohol than medium and 
low quality wines.

Another relationships interesting to evaluate are density - residual sugar and 
density - residual sugar; density decreases while concentration of alcohol 
increases. On the other hand density increases while residual sugar increases. 
The first relation mentiones (density - alcohol) is mores similar to lineal, 
as we can see in the pattern shown in graphs above.

We found a scale that measures the relationship of sugar, acid and pH, which is 
IRF scale. This scale have 4 levels (Dry, Medium Dry, Medium Sweet and Sweet). 
After bivariate analysis, we found that high quality wines evaluated in this 
dataset belong to dry and medium dry, while medium quality wines also have wines 
that belongs to medium sweet.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

We found that there are more concentration of sulphates for medium quality
wine than high quality or low quality. This chemical property acts as 
antinicrobial and antioxidant. So, it seems that is required to be present in 
certain amount. We believed that it seems to have more concentration for medium 
quality wines because there are more wines that belong to this category in this 
dataset.


### What was the strongest relationship you found?
The strongest relationships found are between density-alcohol and 
density-residual sugar

# Multivariate Plots Section

```{r echo=FALSE, warning=FALSE, message=FALSE}
  ggplot(aes( x= alcohol, y = residual.sugar, 
              colour=as.factor(quality)), data = wine) +
  geom_point(position = position_jitter(h=0)) +
  scale_color_brewer(type='seq') +
  facet_wrap(~quality_factor)+
  scale_x_continuous(name = 'Alcohol (% by volume)',
                     lim = c(0, quantile(wine$alcohol, 0.99))) +
  scale_y_continuous(name = 'Residual sugar (g / dm^3)', 
                     lim = c(0, quantile(wine$residual.sugar, 0.99)))+
    ggtitle('Alcohol and residual sugar by quality rating ')
  
```

> Bivariate analysis have showed that there is a correlation between alcohol and
residual sugar. So, would be interesting to evaluated the relation of those 
variables with quality. Graph above shows that wines with more concentration of 
residual sugar are medium quality. For high quality wines you can see that the 
percentage of alcohol increase but the concentration of residual sugar is more 
widespread. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(aes( y= free.sulfur.dioxide/total.sulfur.dioxide, x = alcohol,
            colour=quality_factor), data = wine) +
  
  geom_point(alpha = 0.5, position = position_jitter(h=0)) +
  facet_wrap(~scale)+ 
  scale_color_brewer(type = 'seq', palette = 4) 
 

```


> One relation that we found interesting on bivariate analysis was the relation 
between free.sulfur.dioxide/total.sulfure.dioxide and their impact on quality 
rating. Plot above shows that concentration of SO2 is stronger for medium 
quality wines with a taste dry and medium dry wines, taking into account IRF 
scale. So, it seems that these variables needs to be present in a certain 
concentration, but would have an impact on quality rating if this concentration 
excedees certain amount or is not enough 



```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(aes(x=alcohol, y=density, colour= ratio_scale), data = wine)+
  geom_point()+
  facet_wrap(~quality_factor)+
  scale_x_continuous(name = 'Alcohol (% by volume)', trans = 'sqrt')+
  scale_y_continuous(name = 'Density(g / cm^3)', trans = 'sqrt',
                     limits = c(0.98, 1.01))+
  scale_color_brewer(type = 'div', palette = 3, direction = -1,
                     guide = guide_legend(title = 'IRF scale', 
                    reverse = T, override.aes = list(alpha = 1, size = 2))) +  
  ggtitle('Density vs % of Alcohol by IRF scale and quality rating')
```


> Since alcohol and density are the variable with strongest correlation. 
We found interesting to evaluated how these variables have an impact on quality 
rating taking into account IRF scale. Graph above shows that most of the medium 
sweet wines that belongs to medium quality wines have more density that high 
quality wines, that tends to be medium dry or dry. On the other hand, we can 
see that density for medium dry and dry wines have a similar behaviour for 
medium and high quality wines


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
We found that variable scale created to measure IRF scale have helped to 
evaluate the relationship between my main feautures of interest; such as 
residual sugar, acid and pH. Adding this variable to plot that relates with 
density, alcohol and variables related with SO2 have helped separating by 
quality could help us understand the interactions between them and their impact 
on quality rating. For instance, multivariate analysis has shown us that high 
quality wines tends to be dry and medium dry, with less concentration of SO2 
than medium quality wines and high levels of alcohol percentages than other 
types of wines.


### Were there any interesting or surprising interactions between features?
We found interesting that there are more concentration of SO2 for high quality 
wines. This chemical property is important to be presented on wines for 
microbial prevention and the oxidation of the wine and also affects the nose and 
taste of the wine. So, it would be interested to evaluated this variable with pH 
and residual sugar, which also have an impact on taste in multivariate analysis

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.
No, we didn't create a model


------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE}
ggplot(aes(x= alcohol, colour=quality_factor), data = wine) +
  geom_density()+
  scale_color_brewer(type='seq', palette = 2) +
  labs(x = 'Alcohol (% by volume)', y = 'Density',
          title = 'Distribution of Alcohol by quality')

```

### Description One
> This plot shows density plot for alcohol by categorical variable of quality. 
We found this plot interested to show in this section because it represents the 
distribution of 2 of the variables with strongest correlation. We can notice 
that high quality wine tend to have more % of alcohol by volume than other types 
of whines. Also, shows that there are more concentration of medium quality wines 
with lower percentage of alcohol than low quality wines. However, distribution 
of medium and low quality wines are pretty similar.

### Plot Two
```{r echo=FALSE, warning=FALSE, message=FALSE}
 
ggplot(aes( x= alcohol, y = residual.sugar, 
              colour=as.factor(quality)), data = wine) +
  geom_point(position = position_jitter(h=0)) +
  scale_color_brewer(type='seq') +
  facet_wrap(~quality_factor)+
  scale_x_continuous(name = 'Alcohol (% by volume)',
                     lim = c(0, quantile(wine$alcohol, 0.99))) +
  scale_y_continuous(name = 'Residual sugar (g / dm^3)', 
                     lim = c(0, quantile(wine$residual.sugar, 0.99)))+
    ggtitle('Residual sugar vs % Alcohol by quality rating ')
```



### Description Two

> This plot shows the relation between residual sugar and alcohol on different 
quality of wines. We found this plot usefull because evaluates alcohol variable, 
that has the  strongest relationship with quality and at the same time evaluates 
its relationship with residual sugar (correlated with alcohol) and their impact 
on quality. Also helps to visualize the disribution of different rating in each 
quality category. As mentioned above, you can see how residual sugar are more 
concentrated for medium quality wines and how low quality wines tend to not be 
sweet. 


### Plot Three
```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(aes(x=alcohol, y=density, colour= ratio_scale), data = wine)+
  geom_point()+
  facet_wrap(~quality_factor)+
  scale_x_continuous(name = 'Alcohol (% by volume)', trans = 'sqrt')+
  scale_y_continuous(name = 'Density(g / cm^3)', trans = 'sqrt',
                     limits = c(0.98, 1.01))+
  scale_color_brewer(type = 'div', palette = 3, direction = -1,
                     guide = guide_legend(title = 'IRF scale', 
                    reverse = T, override.aes = list(alpha = 1, size = 2))) +  
  ggtitle('Density vs % of Alcohol by IRF scale and quality rating')
```

### Description Three

> Alcohol and density have the strongest correlation in variables evaluated. 
Plot above shows how these variables have a negative correlation; while 
percentage of alcohol increases density decreases. Also, shows that most of 
medium and high quality wines belong to medium dry and dry scale and that 
density tends to be lower for dry wines than other types of wines.


------

# Reflection

> There are 4898 observations in this dataset that describe quemical properties, 
such as acidity, residual sugar, chlorides, density, pH, sulphates, alcohol and 
quality. In the exploratory analysis we have created new variables; a 
categorical variable for quality named quality_factor with 3 levels(Low, 
Medium and High) and another variables to measure IRF scale; which is an 
industry scale that measures taste based on acid, sugar and pH. This scale is 
created by International Riesling Foundation
(http://drinkriesling.com/tasteprofile/thescale)

> In the first exploratory for univariate analysis, we found that most of wines 
evaluated in this dataset belongs to medium quality. Also, we could notice the 
distribution of main variables of interest.

> For bivariate analysis we've used ggpair in order to detect which variables 
have strongest correlation to evaluated their impact on quality rating. That was
a very usefull tool in order to prioritise variables to be evaluated and their 
relation to each other. We found that variables with strongest correlation were
density-alcohol(-0.795) and density-residual sugar (0.838). Another 
relationships of interest were total.sulfur.dioxide vs free.sulfur.dioxide
(0.597), alcohol vs quality (0.482), alcohol vs residual sugar (0.461), 
total.sulfur.dioxide vs alcohol (0.445), pH vs fixed.acidity (-0.441). 

> We found really usefull the IRF scale; which have helped to understand 
interaction between most important variables evaluated (acid, residual sugar and
pH). Exploratory analysis have shown us that high quality wine tends to have 
more percentage of alcohol, belonged to dry and medium dry type of wine, while
medium quality wines tend to be more sweet with higher concentration of residual 
sugar and less percentage of alcohol and more concentration of SO2.

> Over exploratory analysis project I've struggled to found the best way to 
visualize relationship between some variable and best resolution to show them, 
especially for sulphates and variables related with S02. However, we found 
really interesting to found the relationship between 
free.sulfure.dioxide/total.sulfure.dioxide and quality; which have shown us 
that is necesary to have because it helps to oxidation and taste of the wine, 
but their concentration is least on high quality wine. So, we believed that if 
these concentration excedees certain limits can affect negatively on quality 
rating.

> For future work we could add a model for predict the quality of wine samples, 
or we also could create a model to predict the type of whine based on IRF scale. 
Also, we can create a new dataset combining white and red wine Vinho Verde and 
make a comparison how variables on the dataset differs for each type of whine

